---
title: "TD1"
output: 
  html_document: 
    theme: cerulean
---
## Exercice 1
  1. Create a vector of number
    - The list from 1 to 100 
  
```{r}
seq(1,100)
```
    - A number list/sequence for 10, 20, 25, 50, repeated 5 times ( 10 20 25 50 10 20 25 50 10 20 25 50 10 20 25 50 10 20 25 50)

```{r}
rep(c(10,20,25,50),5)
```
    - a list from 1 to 100, with a step of 5

```{r}
seq(1,100,5)
```
    - Repeat 10 times the number 12

```{r}
rep(12,10)
```
  2. A vector with the number 1, 2, 3 with a repetition of each 4 times with in addition a repetition of the sequence of 4 times
    - Put this vector in the object « VEC »

```{r}
VEC = rep(c(1,2,3),each = 4, times = 4)
VEC
```
    - Multiply all the element of VEC by 5

```{r}
VEC = VEC * 5
VEC
```
    - Calculate the median and the quantile of 75%

```{r}
quantile(VEC, 0.5) # Median
    quantile(VEC, 0.75)
```

  3. Create a vector with a serie of number from 1 to 2000 with a step of 10
    - Put it in the object « vec »

```{r}
vec = seq(1,2000,10)
    vec
```
    - Extract the 10 th value of vec

```{r}
vec[10]
```
    - Display a sub-vector called « vec2 » that corresponds to the values from 2 to 6

```{r}
vec2 = vec[2:6]
    vec2
```
    - Replace the last value of vec2 by 100
    ```{r}
vec2[length(vec2)] = 100
    vec2
```
    - Display vec2 without its 3th value and store it in vec3
    ```{r}
vec3 = vec2[-3]
    vec3
```
    - Replace all the values >= 30 by 30
    ```{r}
vec3[vec3 >= 30] = 30
    vec3
```


## Exercice 2 : Opérations avec les tables
Create a matric with 5 columns and 20 lines with the values from 1 to 100, by lines
  - Call it MAT
```{r}
MAT = matrix(1:100,byrow = TRUE, ncol = 5)
MAT
```
  
  - Display the value at the second line and 5 th column et replace it by NA
```{r}
MAT[2,5] = NA
MAT
```
  - Create “mat “ with only the columns 2,3 and 4 (taking them from MAT)
```{r}
mat = MAT[,2:4]
mat
```
  - Replace the values between 40 and 60 by 50 in the matrix « mat »
```{r}
mat[mat>40 && mat<60] = 50
mat
```
  
## Exercise 3. Operations with a table of decimal numbers
1. Get the file (moodle) M&Ms.xls and open it with excel or open-office
  - save it as MetMs.txt
  - Import it in R and store it in data
```{r}
library(readxl)
data <- read_excel("~/Téléchargements/MetMs.xlsx")
```

2. Create data2 by removing the variable largeur (width) from MetMs and the missing data
```{r}
data2 = subset(data, select = -Forme)
data2 = na.omit(data2)
data2
```

  - Determine for which colors the lengths are minimal and maximal
```{r}
with(data2, Couleur[`Longueur maximale (mm)` == min(`Longueur maximale (mm)`)])
with(data2, Couleur[`Longueur maximale (mm)` == max(`Longueur maximale (mm)`)])
```
  
  - Sort the lines of data2 according to an increasing order of the length
```{r}
data2[order(data2$`Longueur maximale (mm)`),]
```
## Exercice 4. Graphics
```{r}
hist(data2$`Longueur maximale (mm)`, xlab = "Longueur", main = "Longueur des M & M's")
```

```{r}
par(mfrow=c(1,2))
boxplot(data = data2,`Longueur maximale (mm)`~ Couleur , main = "Longueur des M & M's par couleur", xlab= "Couleur", ylab= "Taille")
boxplot(data = data2,`Poids (g)`~ Couleur , main = "Poid des M & M's par couleur", xlab= "Couleur", ylab= "Poid")
```

